---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

This is my submission for Peer Assignment 1.

Ensure that you have the activity.csv file in the RepData_PeerAssessment1 directory,
which should be in your working directory.

Make sure you are using the .csv version, NOT the .zip version (i.e. unzip the file)

## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r}
activity$day <- as.numeric(activity[ , 2], format='%d')
mean_day <- aggregate(activity$steps, by=list(activity$day), FUN=mean, na.action = na.omit)
hist(mean_day[ ,2])
```

## What is the average daily activity pattern?
```{r}
activity_complete <- activity[complete.cases(activity), ]
mean_interval <- aggregate(activity_complete$steps, by=list(activity_complete$interval), FUN=mean, na.action=na.omit)
plot(mean_interval, type = 'l')
```

## Inputing missing values



## Are there differences in activity patterns between weekdays and weekends?
